DNNNNR0: How to Avoid OOTA Without Really Trying

### 2025-03-27 paper/ootavacuous.tex
ootavacuous: Recommend additional standard OOTA advice
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

Because nothing in the C++ memory model rules out such OOTA cycles
(as indicated by the ``sometimes satisfied'' in the figure), the C++
standard explicitly recommends against them in 33.5.4p8
(\co{[atomics.order]})~\cite{ThomasKoeppe2023N4950}:
\begin{quote}
	Implementations should ensure that no “out-of-thin-air” values
	are computed that circularly depend on their own computation.
\end{quote}
This prohibition prevents misapplication of the as-if rule in oracular
C++ implementations.
(The as-if rule permits those optimizations where the optimized code's
observable behavior is the same as if it were unoptimized.)

@@

>>>>

Because nothing in the C++ memory model rules out such OOTA cycles
(as indicated by the ``sometimes satisfied'' in the figure), the C++
standard explicitly recommends against them in 33.5.4p8
(\co{[atomics.order]})~\cite{ThomasKoeppe2023N4950}:
\begin{quote}
	Implementations should ensure that no ``out-of-thin-air'' values
	are computed that circularly depend on their own computation.
\end{quote}
This prohibition prevents misapplication of the as-if rule in oracular
C++ implementations.
(The as-if rule permits those optimizations where the optimized code's
observable behavior is the same as if it were unoptimized.)

@@

\subsection{Possible Change to the Standard}
\label{sec:Possible Change to the Standard}

As noted in Section~\ref{sec:Brief OOTA Overview}, the C++
standard explicitly recommends against them in 33.5.4p8
(\co{[atomics.order]})~\cite{ThomasKoeppe2023N4950}:
\begin{quote}
	Implementations should ensure that no ``out-of-thin-air'' values
	are computed that circularly depend on their own computation.
\end{quote}

Sensible though this recommendation might be, it can unnecessarily raise
anxiety levels of C++ implementers.
After all, what exactly do they need to do in order to follow that
recommendation?
One way to reduce their anxiety levels would be to add something like
the following sentence to this recommendation:
\begin{quote}
	Compiler-based implementations whose binaries run on conventional
	hardware are guaranteed not to compute out-of-thin-air values as
	long as they restrict themselves to thread-at-a-time analysis,
	valid optimizations, and user programs that are free of undefined
	behavior.
\end{quote}
